---
title: "Lesson1: Objects in R"
output: html_notebook
author: "nehaotyagi"
---

# Everything in R is an object

# Functions to identify the object
class() - what kind of object is it (high-level)?
typeof() - what is the object’s data type (low-level)?
length() - how long is it? What about two dimensional objects?
attributes() - does it have any metadata?

# Data Types
    *character*
    *numeric (real or decimal)*
    *integer*

```{r}
x <- "hello"
y <- 2.45
z <- 2L

cat("class of x is:",class(x), "; typeof x is:", typeof(x), "\n")
cat("class of y is:",class(y), "; typeof y is:", typeof(y), "\n")
cat("class of z is:",class(z), "; typeof z is:", typeof(z), "\n")

```

*logical*
*complex*

```{r}
a <- y > z
b <- 1-2i

cat("class of a is:",class(a), "; class of b is:",class(b))

```

# Data Structures
    Homogeneous
      1D: Atomic Vector; 2D: Matrix; 3D: Array
    Heterogeneous
      1D: List; 2D: Data Frame
    
*Vectors*
    *atomic vector* - all elements of an atomic vector must be the same type
    *list* - elements of a list can have different types

```{r}
w <- c(1,2,3,4)
x <- "hello"
y <- c("hello", 2L)
z <- list("hello", 2L)

cat("w is an atomic vector:", is.atomic(w), " w is a list:", is.list(w), "\n")
cat("x is an atomic vector:", is.atomic(y), " x is a list:", is.list(y), "\n")
cat("y is an atomic vector:", is.atomic(y), " y is a list:", is.list(y), "\n")
cat("z is an atomic vector:", is.atomic(z), " z is a list:", is.list(z), "\n")

# Note
a <- 2
cat("a is an atomic vector:", is.atomic(a), " length of a is:", length(a), "\n")

a[3] <- 4 # Can I add elements to this vector a?
a
cat("a is an atomic vector:", is.atomic(a), " length of a is:", length(a), "\n")

a[0]    # Does index 0 exist?
```
# Observations: 
is.vector() returns TRUE if the object is a vector with no attributes other than names. 
is.atomic() or is.list can be used to test the type of the vector
A scalar is stored in R as a vector of length 1.  It allows us to add elements to it
Indexing in R begins at 1; numeric(0) indicates vector of length 0 i.e. empty

# Vector recycling
    If R encounters two vectors of different lengths in a binary operation, 
    it replicates (recycles) the smaller vector until it is the same length as the longest vector 
    and then it does the operation.
```{r}
a <- 1:3
b <- 2:5
a
b
a + b
```

*matrix*
```{r}

```

*data frame*
```{r}

```

*factors*
```{r}

```
# Attributes in R
    
# Classes in R

What is a class?  
In object-oriented programming, a class is a blueprint for creating objects (a particular data structure), providing initial values for state (member variables or attributes), and implementations of behavior (member functions or methods). (https://brilliant.org/wiki/classes-oop/)

# S3 class

# S2 class

# Reference class