---
title: "Lesson1: Objects in R"
output: html_notebook
author: "nehaotyagi"
---

# Everything in R is an object

# Functions to identify the object
class() - what kind of object is it (high-level)?
typeof() - what is the object’s data type (low-level)?
length() - how long is it? What about two dimensional objects?
attributes() - does it have any metadata?

# Data Types
    *character*
    *numeric (real or decimal)*
    *integer*

```{r}
x <- "hello"
y <- 2.45
z <- 2L

cat("class of x is ",class(x))
class(y)
class(z)
```
    *logical*
    *complex*
```{r}
a <- y > z
b <- 1-2i

cat("class of a is ",class(a), ": class of b is ",class(b))

```

# Data Structures
    *atomic vector*
    *list*
    *matrix*
    *data frame*
    *factors*



# Classes in R

What is a class?  
In object-oriented programming, a class is a blueprint for creating objects (a particular data structure), providing initial values for state (member variables or attributes), and implementations of behavior (member functions or methods). (https://brilliant.org/wiki/classes-oop/)

# S3 class

# S2 class

# Reference class